### 2021-10-27 documentos/proofdiagonalization.tex
Choi_Paulis

\section{Diagonalization of Choi Matrix $\mathcal{D}_N$} % {{{
\label{sec:appendix:diagonalization}
% 
\subsection{Single qubit}
In order to simplify the derivation of the relations, let us employ pairs of binary indices instead of a single quaternary, i.e. $\alpha\to\mu+2\nu$. Hence we can write the elements of the Pauli basis
$(\sigma_{0},\sigma_{1},i\sigma_{2},\sigma_{3})$, compactly as
$\sigma_{\mu\nu}=\sum_{j=0}^{1} (-1)^{j \mu}\dyad{j}{j\oplus \nu}$. In vectorized form
\begin{equation}
\kket{\sigma_{\mu\nu}}=\sum_{j=0}^{1} (-1)^{j \mu}\kket{j,j\oplus \nu},
\end{equation}
in the same way,
\begin{equation}
\kket{l,l\oplus n}=\frac{1}{2}\sum_{m=0}^{1} (-1)^{ml}\kket{\sigma_{mn}}.
\label{PauToComp}
\end{equation}
On the other hand, the matrix form of an arbitrary Pauli map $\mathcal{E}$ may be written as
\begin{align}
\hat{\mathcal{E}} & = \frac{1}{2}\sum_{\mu\nu=0}^{1} \tau_{\mu\nu} \ddyada{\sigma_{\mu\nu}},\\
& = \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} \ddyad{j,j\oplus\nu}{k,k\oplus\nu}.
% \label{Map_Pauli}
\end{align}
After applying the reshuffling operation on $\hat{\mathcal{E}}$, we obtain the Choi matrix associated to the map. It reads
\begin{equation}
 \mathcal{D} = 
  \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} 
      \ddyad{j,k}{j\oplus\nu,k\oplus\nu}.
\label{Choi_a1}
\end{equation}

Let us apply the index relabeling $k\to j\oplus m$, then the Choi matrix holds
\begin{equation}
 \mathcal{D} =
    \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu} (-1)^{\mu m} 
        \ddyad{j,j\oplus m}{j\oplus\nu,j\oplus\nu\oplus m}.
\label{Choi_a1}
\end{equation}
Due to the fact that we are dealing with binary indices, thus we can safely write the modular sums in the exponents as standard sums. By employing the relation between computational and Pauli elements (Eq. \ref{PauToComp}), by noticing that $\sum_j(-1)^{j(m\pm n)}=2\delta_{mn}$, and after some steps we have
\begin{align*}
 \mathcal{D} & = 
%  \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu}(-1)^{\mu m}  \left( \frac{1}{2}\sum_{k=0}^{1} (-1)^{kj}\kket{\sigma_{km}}\right)\left(\frac{1}{2}\sum_{n=0}^{1} (-1)^{n(j\oplus\nu)}\bbra{\sigma_{nm}}\right),\\
\frac{1}{8}\sum_{jkmn\mu\nu} \tau_{\mu\nu}(-1)^{\mu m+kj+n(j\oplus\nu)}  \ddyad{\sigma_{km}}{\sigma_{nm}},\\
%  & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \left(\sum_{j} (-1)^{j(k+n)} \right) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
%  & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} (2\delta_{kn}) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
 & = \frac{1}{4}\sum_{mn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \ddyad{\sigma_{nm}}{\sigma_{nm}},\\
 & = \frac{1}{2}\sum_{mn} \left(\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu}\right) \ddyad{\sigma_{nm}}{\sigma_{nm}},
% \label{Choi_a1}
\end{align*}
Then $\mathcal{D}$ is diagonal in the Pauli basis, with eigenvalues given by
\begin{equation}
\lambda_{mn}=\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu},
\end{equation}
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes
H$ and $H$ is the Hadamard matrix.

>>>>

\section{Diagonalization of Choi Matrix $\mathcal{D}_N$} % {{{
\label{sec:appendix:diagonalization}
% 
\subsection{Single qubit}
In order to simplify the derivation of the relations, let us employ pairs of binary indices instead of a single quaternary, i.e. $\alpha\to\mu+2\nu$. Hence we can write the elements of the Pauli basis
$(\sigma_{0},\sigma_{1},i\sigma_{2},\sigma_{3})$, compactly as
$\sigma_{\mu\nu}=\sum_{j=0}^{1} (-1)^{j \mu}\dyad{j}{j\oplus \nu}$. In vectorized form
\begin{equation}
\kket{\sigma_{\mu\nu}}=\sum_{j=0}^{1} (-1)^{j \mu}\kket{j,j\oplus \nu},
\end{equation}
in the same way,
\begin{equation}
\kket{l,l\oplus n}=\frac{1}{2}\sum_{m=0}^{1} (-1)^{ml}\kket{\sigma_{mn}}.
\label{PauToComp}
\end{equation}
On the other hand, the matrix form of an arbitrary Pauli map $\mathcal{E}$ may be written as
\begin{align}
\hat{\mathcal{E}} & = \frac{1}{2}\sum_{\mu\nu=0}^{1} \tau_{\mu\nu} \ddyada{\sigma_{\mu\nu}},\\
& = \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} \ddyad{j,j\oplus\nu}{k,k\oplus\nu}.
% \label{Map_Pauli}
\end{align}
After applying the reshuffling operation on $\hat{\mathcal{E}}$, we obtain the Choi matrix associated to the map. It reads
\begin{equation}
 \mathcal{D} = 
  \frac{1}{2}\sum_{jk\mu\nu} \tau_{\mu\nu} (-1)^{\mu(j+k)} 
      \ddyad{j,k}{j\oplus\nu,k\oplus\nu}.
\label{Choi_a1}
\end{equation}
Note furthermore that the expression above may also be written as a combination of tensor products of Pauli matrices
% 
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} \sigma_{\mu\nu} \otimes \sigma_{\mu\nu}^*.
 \label{Choi_Pauli_1q}
\end{equation}
% 
Back to the Eq. \ref{Choi_a1}), let us apply the index relabeling $k\to j\oplus m$, then the Choi matrix holds
\begin{equation}
 \mathcal{D} =
    \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu} (-1)^{\mu m} 
        \ddyad{j,j\oplus m}{j\oplus\nu,j\oplus\nu\oplus m}.
% \label{Choi_a1}
\end{equation}
Due to the fact that we are dealing with binary indices, thus we can safely write the modular sums in the exponents as standard sums. By employing the relation between computational and Pauli elements (Eq. \ref{PauToComp}), by noticing that $\sum_j(-1)^{j(m\pm n)}=2\delta_{mn}$, and after some steps we have
\begin{align*}
 \mathcal{D} & = 
%  \frac{1}{2}\sum_{jm\mu\nu} \tau_{\mu\nu}(-1)^{\mu m}  \left( \frac{1}{2}\sum_{k=0}^{1} (-1)^{kj}\kket{\sigma_{km}}\right)\left(\frac{1}{2}\sum_{n=0}^{1} (-1)^{n(j\oplus\nu)}\bbra{\sigma_{nm}}\right),\\
\frac{1}{8}\sum_{jkmn\mu\nu} \tau_{\mu\nu}(-1)^{\mu m+kj+n(j\oplus\nu)}  \ddyad{\sigma_{km}}{\sigma_{nm}},\\
%  & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \left(\sum_{j} (-1)^{j(k+n)} \right) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
%  & = \frac{1}{8}\sum_{kmn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} (2\delta_{kn}) \ddyad{\sigma_{km}}{\sigma_{nm}},\\
 & = \frac{1}{4}\sum_{mn\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu} \ddyad{\sigma_{nm}}{\sigma_{nm}},\\
 & = \frac{1}{2}\sum_{mn} \left(\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu}\right) \ddyad{\sigma_{nm}}{\sigma_{nm}},
% \label{Choi_a1}
\end{align*}
Then $\mathcal{D}$ is diagonal in the Pauli basis, with eigenvalues given by
\begin{equation}
\lambda_{mn}=\frac{1}{2}\sum_{\mu\nu} \tau_{\mu\nu} (-1)^{\mu m+n\nu},
\end{equation}
or more compactly $\vec{\lambda} = \sa~\vec{\tau}$, where $\sa=2^{-1}H\otimes
H$ and $H$ is the Hadamard matrix.
